---
title: "Assignment 9: Mapping"
author: "Student Name"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Water Data Analytics on mapping

## Directions
1. Change "Student Name" on line 3 (above) with your name.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, check your PDF against the key and then submit your assignment completion survey at https://forms.gle/NDWEUu73LooFJPVM8 

Having trouble? See the assignment's answer key if you need a hint. Please try to complete the assignment without the key as much as possible - this is where the learning happens!

Target due date: 2022-04-12

## Setup

1. Load the tidyverse, LAGOSNE, maps, dataRetrieval, sf, and nhdplusTools packages. Set your ggplot theme (can be theme_classic or something else).


```{r setup}
library(tidyverse)
library(dataRetrieval)
library(maps)
library(sf)
install.packages("nhdplusTools")
#library(nhdplusTools)

theme_set(theme_classic())
options(scipen = 5)
```

## LAGOS-NE

2. Choose five lakes in the LAGOS-NE database that are located within the same state and have chlorophyll data. Subset your data accordingly, and create two plots:

* A map of the five lakes within the state boundary, with each lake point as its own color.
* A boxplot with jittered points layered on top of chlorophyll concentrations in each lake (chlorophyll on y axis, lake on x axis), with each lake having a fill and/or color that matches the map. 

```{r}
LAGOSdatabase <- lagosne_load()

Mystate <- st_as_sf(map(database = "state", plot = FALSE,
                        fill = TRUE))

Iowa_subset <- filter(Mystate, ID == "iowa")

Lagoslocus <- LAGOSdatabase$locus
LAGOSstate <- LAGOSdatabase$state
LAGOSnutrient <- LAGOSdatabase$epi_nutr

LAGOS_Together <- left_join(LAGOSnutrient, Lagoslocus) %>% left_join(., LAGOSstate) %>% 
  filter(state == "IA") %>% 
  select(lagoslakeid, sampledate, chla, nhd_lat, nhd_long) %>%
drop_na(chla) %>%
arrange(lagoslakeid)
 
lake_sample <- sample(LAGOS_Together$lagoslakeid, size = 5)

LAGOSsample <- LAGOS_Together %>% 
  filter(lagoslakeid %in% lake_sample)
colnames(LAGOSsample)

LAGOSsample_sp <- st_as_sf(LAGOSsample, coords = c("nhd_long", "nhd_lat"), crs = 4326)

ggplot() +
  geom_sf(data = Iowa_subset, fill = "white") +
  geom_sf(data = LAGOSsample_sp, aes(color =as.factor(lagoslakeid)))+
scale_color_viridis_d() +
labs(color = "Lake ID") +
theme(legend.position = "top")

ggplot(LAGOSsample, aes(as.factor(lagoslakeid), y = chla, fill = as.factor(lagoslakeid))) +
  geom_boxplot() +
  geom_jitter()+
  scale_y_log10() +
scale_fill_viridis_d() +
labs(x = "Lake ID", y = expression("Chlorophyll a ("*mu*"g/L)"), fill = "") +
theme(legend.position = "none")


```


## NHDPlus

3. Delineate and map the watershed and flowpaths in the network upstream of the Eno River gage in Durham (USGS-02085070). Your map should include all stream orders.


```{r}
EnoParams <- whatNWISdata(siteNumbers = "02085070")
EnoCoords <- EnoParams %>%
select(site_no, dec_lat_va, dec_long_va) %>%
distinct()

start_point <- st_sfc(st_point(c(EnoCoords$dec_long_va, EnoCoords$dec_lat_va)),
crs = 4269)

start_comid <- discover_nhdplus_id(start_point)

NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid),
mode = "upstreamTributaries",
distance_km = 1000)

subset_file <- tempfile(fileext = ".gpkg")
subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid),
output_file = subset_file,
nhdplus_data = "download",
flowline_only = FALSE,
return_data = TRUE, overwrite = TRUE)

flowline <- subset$NHDFlowline_Network
catchment <- subset$CatchmentSP

plot_nhdplus("USGS-02085070")

max(flowline$totdasqkm)
summary(as.factor(flowline$streamorde))

gages <- get_nwis(AOI = catchment)
```

4. What is the upstream drainage area for this gage, and how are the flowpaths distributed with regard to stream order?

> 368 sq kilometers with 123, 45, 19, and 28 flowpaths of respective stream order of 1, 2,3,and 4

5. Are there any other gage sites in the upstream catchment? If so, where are they?
>  we have cole mill road and hillsborough 
